\chapter{Brevity}
\label{brevity}

Like most professions, combat aviation has its own jargon.
Called \emph{brevity}, it's designed to let pilots communicate
as quickly as possible in the heat of battle.

Most folks you find online will be using \textsc{us/nato} brevity.
There's many,
many terms,\punckern\footnote{An early-2000s version of an official
multiservice \textsc{us} training guide is declassified and freely available
online at \url{https://apps.dtic.mil/dtic/tr/fulltext/u2/a404426.pdf}}
but we'll cover some of the most common ones.
Before we dive in, though, a few points:
\begin{enumerate}
\item Basic radio protocol is, ``<you>, <me>. <what I have to say>''
    For example, if my callsign is Warhawk and I want to ask another flight,
    callsign Hammer, what they're up to, I'd say,
    ``Hammer, Warhawk. Status''

    Of course, this isn't an absolute rule.
    If I'm leading a flight and I want us to get ready for combat, I'll say,
    ``Flight, fence in''\quotekern, not, ``Flight, lead. Fence in.''
    Similarly, my wingman will just say, ``Two is fenced in\dots''
    not, ``Two, lead, fenced in\dots''
    Don't be an annoying pedant.

\item If you can't think of the right brevity,
    \emph{use plain language!}
    The main goal is to be understood, quickly.
    Sounding cool is an added bonus.

\item Provide additional information whenever you can.
    If you're shooting an AMRAAM, ``Fox~3'' isn't a very helpful call---it tells
    me a missile is coming off your jet, but it could be pointed at my ass
    for all I know.
    ``Fox~3, Bullseye 069 42'' lets me know what the intended target is,
    and I'll maneuver (and engage other targets!) to best support you.
    Similarly, if the two of us merge with a pair of bandits,
    and you call ``Fox 2, \designation{Mig-21} low''\quotekern,
    I'll take the high one.

    Work with friendlies to build a shared unerstanding of what's going on
    around you.

\item Some terms have one version for friendly forces
    and another for hostiles or unknown forces.
    Don't mix them up!
    If you're \emph{visual} on your wingman,
    you can fly in formation with them.
    But if you \emph{tally} your wingman,
    they'll wonder why you're trying to shoot them.
\end{enumerate}

\section{Who? Where?}

\begin{description}
\item[BRAA] Short for \emph{Bearing, Range, Altitude, Aspect}.
    From my plane, the thing I'm describing is in this direction
    (bearing), for this distance (range), at this altitude.
    Provide the direction they're moving (aspect) if possible.

    ``Bogey, BRAA 035 50, 18 thousand, hot.''

    Bearings are in degrees and ranges are in miles.
    Altitudes are in thousands of feet, called ``thousand'' for hostiles
    and ``angels'' for friendlies. Aspects include:
    \begin{description}
    \item[Hot] heading towards you
    \item[Dragging] heading away from you
    \item[Beaming] heading perpendicular to you (provide a direction)
    \end{description}

\item[Bullseye] The thing I'm describing is this direction and distance
    from the \emph{bullseye}.
    The bullseye is a preplanned location everyone has agreed to use for
    a reference. It could be a large ground target (like an airfield),
    some landmark, or just an arbitrary point---all that matters is
    everyone knows where it is.

    ``Engaging armored column, Bullseye 180 20''

\item[Capture] I see a ground target on my sensors (such as radar
    or targeting pod).

\item[Contact] I see a potential target on my sensors (usually radar).
    Also brevity for the potential target itself.
    Follow with a location.

    ``Contact, Bullseye 120 35, 18 thousand. Hot.''

\item[Declare] Tell me if the contact at the given location
    (using BRAA or a Bullseye call) is friendly or not (see below).
    Requested from command and control \acronym{(c2)} assets,
    like an airborne \designation{E-3} Sentry
    or a ground-based fighter controller.

\item[Friendly/Bogey/Outlaw/Bandit/Hostile]
    Air contacts, by increasing threat:
    \begin{description}
    \item[Friendly] An positively-identified allied contact.
    \item[Bogey] An unknown contact.
    \item[Outlaw] A bogey that is likely a hostile
        (e.g., that launched from an enemy airfield)
    \item[Bandit] An enemy contact---does not imply permission to engage.
    \item[Hostile] An enemy contact---free to fire.
    \end{description}

\item[Furball] A dogfight of friendlies and hostiles.
    Don't shoot an active radar missile into these or it might go pitbull
    on a friendly.

\item[Group] Air contacts flying together.

\item[Merged] Contacts---usually bandits---are within visual range
    \acronym{(wvr)}.
    Get your eyes out of the cockpit and start turning \& burning.
    Named after what happens to multiple contacts when they run into each other
    on someone's radar screen---they \emph{merge}.

\item[Picture] Tell me about nearby contacts (with BRAA or Bullseye calls).
    Requested from \acronym{c2} assets.

\item[Pop-up] A new contact has suddenly appeared at the given location.

\item[Tally/No Joy] I see the enemy.\,/\,I don't see the enemy.

\item[Tumbleweed] I've lost situational awareness \acronym{(sa)}
    and don't know what's going on.
    Help me get reoriented, if you can.

\item[Visual/Blind] I see the friendly.\,/\,I don't see the friendly.
\end{description}

\section{Administration}

What are you doing? What's your status?
\begin{description}

\item[As Fragged] I'm continuing as-planned.

\item[Bingo] I'm low on fuel and need to leave right now if I want enough
    to land.

\item[Commit] Intercept the given group.

    You usually brief a \emph{commit range}---if contacts get within that range,
    you stop what you're doing and commit on those contacts.

\item[Fence In] Get ready for combat---shut off external lights,
    arm weapons, enable electronic countermeasures \acronym{(ecm)}, etc.
    From ``crossing the fence'' as you enter bad guy territory.
    Once you're out of combat, you \emph{fence out}.

\item[Gate/Buster] Fly as fast as you can [with/without] afterburner.

\item[Joker] I'm starting to run low on fuel and should think about getting
    out of here. Comes before bingo.

\item[Rolex] Delay the (package's) plan by the given amount of time.

    ``Package, rolex three minutes.''
\end{description}

\section{Combat}

Shooting things and not getting shot.
\begin{description}

\item[Defending/Engaged defensive] I'm busy maneuvering to avoid getting shot by
    a missile. Provide additional information if you can.

    ``Defending \designation{SA-5}, West!''

\item[Delouse] A hostile jet is chasing me; please shoot them!

\item[Dirt] A \acronym{sam} radar appeared on my \acronym{rwr}.

\item[Fox 1] I'm shooting a semi-active radar homing \acronym{(sarh)}
    missile, like an \designation{AIM-7} Sparrow.
\item[Fox 2] I'm shooting a heat-seeking missile, like an
    \designation{AIM-9} Sidewinder.
\item[Fox 3] I'm shooting an active radar homing \acronym{(arh)} missile,
    like an \designation{AIM-120} AMRAAM.

\item[Guns] I'm shooting my cannon.

\item[Nails] An aircraft's search radar appeared on my \acronym{rwr}.

\item[Magnum] I'm shooting an anti-radiation missile,
    like an \designation{AGM-88} HARM.

\item[Paveway] I'm dropping a laser-guided bombs,
    like a \designation{GBU-}\designation{10/12/16}

\item[Pickle] I'm dropping bombs (usually unguided),
    like the \designation{Mark 80} series.

\item[Pig] I'm dropping a glide weapon, like a \designation{AGM-154} JSOW.

\item[Rifle] I'm shooting an air-to-ground missile,
    like an \designation{AGM-65} Maverick.

\item[SAM] I \emph{see} an incoming surface to air missile!

\item[Singer] My \acronym{rwr} indicates a \acronym{sam} is shooting at me!

\item[Slapshot] Shoot an anti-radiation missile at the given radar/direction \\
    (as fast as possible, because \emph{they're shooting at me!})

    ``Slapshot \designation{SA-3}, South!''

\item[Sniper] Shoot an anti-radiation missile at the given \acronym{sam} radar.

    ``Sniper \designation{SA-2}, Bullseye 270 30.''

\item[Spike/Mud Spike] An [aircraft/\acronym{sam}] radar is tracking me. \\
    (Now might be a good idea to start defending.)

\item[Splash] Target destroyed. Sometimes \textbf{Shack} is used for
    air-to-ground impacts.

\end{description}
